\chapter{Introduction}
\label{chap:introduction}

\section{Background}
\label{section:background}

Figure skating attracts millions of viewers during Winter Olympics and ISU Grand Prix events. The sport features six primary jumps: toe loop, flip, and Lutz use toe-assisted takeoff, while Salchow, loop, and Axel use edge takeoff only. These jumps differ by subtle takeoff mechanics that challenge even experienced viewers. Axel jumps take off forward-facing and require about 0.5 extra rotation compared to other jumps. During live competitions, technical panels review unclear calls using instant replays. Broadcasters typically read from pre-written program sheets, which often misidentify jumps and leave casual audiences confused about scoring.

\section{Problem Statement}
\label{section:problem-statement}

Figure skating jumps look similar but receive different technical scores. Current solutions create several problems for viewers:

\begin{itemize}
\item Manual review by technical panels slows competition and remains inaccessible to broadcast audiences
\item Broadcasters read program sheets without deep jump analysis and frequently confuse similar jumps
\item Casual fans cannot connect visual jump mechanics to resulting scores
\item Even experts disagree on subtle edge calls and rotation counts during live events
\end{itemize}

Traditional rule-based computer vision fails on variable skater styles and camera angles. This project uses AI to learn jump patterns directly from labeled video data.

\section{Solution Overview: FigureFlow}
\label{section:solution-overview}

FigureFlow provides real-time jump identification through a three-stage computer vision pipeline:

\begin{enumerate}
\item YOLOv8 detects and tracks skaters across video frames from broadcast footage
\item MediaPipe Pose extracts 33 body keypoints per frame (x, y, z coordinates plus visibility) 
\item LSTM neural network analyzes 100-frame pose sequences to identify takeoff patterns and classify jumps
\end{enumerate}

The system delivers results in two phases:
\begin{itemize}
\item Phase 1 (baseline): Axel vs non-Axel classification 
\item Phase 2 (extended): Toe jumps vs edge jumps classification 
\end{itemize}

Data comes from YouTube competition clips labeled using ISU Judges Details PDFs plus the FS-Jump3D dataset containing 253 multi-view videos with 3D pose ground truth.

\section{Target Users}
\label{section:target-user}

FigureFlow serves three main groups:

\textbf{Casual viewers} represent the primary audience. These are sports fans watching Olympics or Grand Prix events who want simple jump explanations without technical jargon. They need visual overlays showing jump names and basic mechanics.

\textbf{Serious enthusiasts} follow ISU competitions closely. They want verification of technical panel calls with precision above 87\% to build trust in the system.

\textbf{Broadcasters and online platforms} need reliable jump identification to replace error-prone program sheets. The system requires low latency for integration with live video streams.

\section{Terminology}
\label{section:terminology}

\begin{description}
\item[Toe Jump] Three jumps using toe pick assistance: toe loop, flip, and Lutz
\item[Edge Jump] Three jumps using only blade edge: Salchow, loop, and Axel  
\item[Axel Jump] Forward-facing takeoff requiring 2.5x or 3.5x rotations for double/triple
\item[Pose Keypoints] 33 MediaPipe landmarks tracking hips, knees, ankles, shoulders
\item[Temporal Sequence] 100-frame windows capturing entry, takeoff, flight, and landing phases
\item[Technical Panel] ISU officials who identify jumps live using video replay
\end{description}

\section{Technical Requirements}
\label{section:requirements}

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|}
\hline
\textbf{Task} & \textbf{Clips Required} \\
\hline
Axel vs Non-Axel & 200--300 \\
Toe vs Edge & 400--600 \\
Six Jumps (exploratory) & 600--1000 \\
\hline
\end{tabular}
\caption{Data requirements by development phase}
\end{table}

FigureFlow beats human judging error rates while remaining accessible to non-expert audiences.
